---
title: "2026-04-20_hw08"
author: "Chris"
date: "2026-04-20"
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    theme: cosmo
    css: style.css
---


# Read the Data

```{r message=FALSE, warning=FALSE}
library(readxl)
library(tidyverse)
library(lmerTest)
library(broom.mixed)

df_health <- read_excel("data/HW8-HEALTH.xlsx") |>
  janitor::clean_names() |>
  mutate(time = factor(time))
```


# Question a

## Create a 2 by 2 Table

```{r}
df_baseline <- df_health |>
  filter(time == "1")

tab_baseline <- table(df_baseline$txt, df_baseline$health)
tab_baseline
round(prop.table(tab_baseline, margin = 1), 3)
```

## Chi-Square Test

```{r}
chisq.test(tab_baseline)
```


## Fisher's Exact Test

```{r}
fisher.test(tab_baseline)
```


* At the time of randomization, 48.8% of participants in the control group rated their health as good, compared with 41.0% in the intervention group. Conversely, 51.2% of the control group and 59.0% of the intervention group rated their health as poor. Although the intervention group appeared to have a slightly higher proportion of poor self-rated health at baseline, the difference was small.  

* A Pearson chi-square test showed no statistically significant association between randomized group assignment and baseline health self-rating ($\chi^2 = 0.223, df = 1, p = 0.637 > 0.05$). Fisher’s exact test gave a similar conclusion ($p = 0.509$). Thus, there is no evidence that self-rated health differed by treatment group at the time of randomization.  

* These findings suggest that the two randomized groups were reasonably comparable with respect to baseline health self-rating, which is consistent with what we would expect from a randomized trial. Therefore, any differences observed during follow-up are less likely to be explained by baseline imbalance in this variable.  







































