---
title: "2026-04-20_hw08"
author: "Chris"
date: "2026-04-20"
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    theme: cosmo
    css: style.css
---

The R project is on GitHub: https://github.com/ChrisW12372/2026-04-20_p8131-hw08.git

# Read the Data

```{r message=FALSE, warning=FALSE}
library(readxl)
library(tidyverse)
library(lmerTest)
library(broom.mixed)
library(geepack)

df_health <- read_excel("data/HW8-HEALTH.xlsx") |>
  janitor::clean_names() |>
  mutate(
    id = factor(id),
    time = factor(time),
    txt = factor(txt),
    health = factor(health),
    agegroup = factor(agegroup)
    )
```


# Question a

## Create a 2 by 2 Table

```{r}
df_baseline <- df_health |>
  filter(time == "1")

tab_baseline <- table(df_baseline$txt, df_baseline$health)
tab_baseline
round(prop.table(tab_baseline, margin = 1), 3)
```

## Chi-Square Test

```{r}
chisq.test(tab_baseline)
```


## Fisher's Exact Test

```{r}
fisher.test(tab_baseline)
```


* At the time of randomization, 48.8% of participants in the control group rated their health as good, compared with 41.0% in the intervention group. Conversely, 51.2% of the control group and 59.0% of the intervention group rated their health as poor. Although the intervention group appeared to have a slightly higher proportion of poor self-rated health at baseline, the difference was small.  

* A Pearson chi-square test showed no statistically significant association between randomized group assignment and baseline health self-rating ($\chi^2 = 0.223, df = 1, p = 0.637 > 0.05$). Fisher’s exact test gave a similar conclusion ($p = 0.509$). Thus, there is no evidence that self-rated health differed by treatment group at the time of randomization.  

* These findings suggest that the two randomized groups were reasonably comparable with respect to baseline health self-rating, which is consistent with what we would expect from a randomized trial. Therefore, any differences observed during follow-up are less likely to be explained by baseline imbalance in this variable.  


# Question b

## Some Preliminary Procedures

```{r}
df_baseline <- df_baseline |>
  select(id, baseline_health = health)

df_followup <- df_health |>
  filter(time != "1")

df_gee <- df_followup |>
  left_join(df_baseline, by = "id") |>
  mutate(
    health_bin = if_else(health == "Good", 1, 0),
    baseline_bin = if_else(baseline_health == "Good", 1, 0),
    txt = relevel(txt, ref = "Control"),
    time = factor(time, levels = c("2", "3", "4")),
    agegroup = relevel(agegroup, ref = "15-24")
  )
```


## Fit the GEE Models

```{r}
fit_gee <- geeglm(
  health_bin ~ baseline_bin + txt + time + agegroup, id = id,
  data = df_gee, 
  family = binomial(link = "logit"),
  corstr = "unstructured"
)

summary(fit_gee)
```


## Get Odds Ratios

```{r}
coef_gee <- coef(summary(fit_gee))

tab_or_gee <- tibble(
  term = rownames(coef_gee),
  estimate = coef_gee[, "Estimate"],
  se = coef_gee[, "Std.err"],
  p_value = coef_gee[, "Pr(>|W|)"]
) |>
  mutate(
    OR = exp(estimate),
    lower_95 = exp(estimate - 1.96 * se),
    upper_95 = exp(estimate + 1.96 * se)
  )

tab_or_gee
```


* A logistic GEE model with an unstructured working correlation matrix was fit to the repeated follow-up observations, excluding the randomization visit. The outcome was participants’ self-rated health during follow-up, and the predictors were baseline health status, treatment group, follow-up visit and age group. This model accounts for the correlation among repeated observations from the same participant.  

* Baseline health was a strong predictor of follow-up self-rated health. After adjustment for treatment group, follow-up visit, and age group, participants who reported good health at baseline had significantly higher odds of reporting good health during follow-up than those who reported poor health at baseline (OR = 5.83, 95% CI: 2.07–16.45, p < 0.001).   

* Treatment group was also significantly associated with follow-up health. Participants in the intervention group had substantially higher odds of reporting good health during follow-up than those in the control group (OR = 6.75, 95% CI: 2.35–19.45, p < 0.001), after controlling for baseline health, follow-up visit, and age group.  

* There was no statistically significant evidence that self-rated health differed across follow-up visits. Relative to `time = 2`, the odds of reporting good health were not significantly different at `time = 3` (OR = 1.31, 95% CI: 0.67–2.56, p = 0.422) or `time = 4` (OR = 1.36, 95% CI: 0.57–3.21, p = 0.489).  

* Age group was partially associated with follow-up health. Compared with participants aged 15–24 years, those aged 25–34 years had significantly higher odds of reporting good health during follow-up (OR = 5.21, 95% CI: 1.87–14.48, p = 0.002). The estimated odds were also higher for participants aged 35 years and older, but this difference was not statistically significant (OR = 2.72, 95% CI: 0.46–16.07, p = 0.268).  

* Overall, these results suggest that follow-up self-rated health was strongly associated with baseline health and treatment assignment, while there was no clear evidence of a significant difference across follow-up visits. The results also suggest a higher likelihood of good self-rated health among participants aged 25–34 years compared with those aged 15–24 years.  






























