---
title: "2026-04-20_hw08"
author: "Chris"
date: "2026-04-20"
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    theme: cosmo
    css: style.css
---

The R project is on GitHub: https://github.com/ChrisW12372/2026-04-20_p8131-hw08.git

# Read the Data

```{r message=FALSE, warning=FALSE}
library(readxl)
library(tidyverse)
library(broom.mixed)
library(geepack)
library(lme4)

df_health <- read_excel("data/HW8-HEALTH.xlsx") |>
  janitor::clean_names() |>
  mutate(
    id = factor(id),
    time = factor(time),
    txt = factor(txt),
    health = factor(health),
    agegroup = factor(agegroup)
    )
```


# Question a

## Create a 2 by 2 Table

```{r}
df_baseline <- df_health |>
  filter(time == "1")

tab_baseline <- table(df_baseline$txt, df_baseline$health)
tab_baseline
round(prop.table(tab_baseline, margin = 1), 3)
```

## Chi-Square Test

```{r}
chisq.test(tab_baseline)
```


## Fisher's Exact Test

```{r}
fisher.test(tab_baseline)
```


* At the time of randomization, 48.8% of participants in the control group rated their health as good, compared with 41.0% in the intervention group. Conversely, 51.2% of the control group and 59.0% of the intervention group rated their health as poor. Although the intervention group appeared to have a slightly higher proportion of poor self-rated health at baseline, the difference was small.  

* A Pearson chi-square test showed no statistically significant association between randomized group assignment and baseline health self-rating ($\chi^2 = 0.223, df = 1, p = 0.637 > 0.05$). Fisher’s exact test gave a similar conclusion ($p = 0.509$). Thus, there is no evidence that self-rated health differed by treatment group at the time of randomization.  

* These findings suggest that the two randomized groups were reasonably comparable with respect to baseline health self-rating, which is consistent with what we would expect from a randomized trial. Therefore, any differences observed during follow-up are less likely to be explained by baseline imbalance in this variable.  


# Question b

## Some Preliminary Procedures

```{r}
df_baseline <- df_baseline |>
  select(id, baseline_health = health)

df_followup <- df_health |>
  filter(time != "1")

df_gee <- df_followup |>
  left_join(df_baseline, by = "id") |>
  mutate(
    health_bin = if_else(health == "Good", 1, 0),
    baseline_bin = if_else(baseline_health == "Good", 1, 0),
    txt = relevel(txt, ref = "Control"),
    time = factor(time, levels = c("2", "3", "4")),
    agegroup = relevel(agegroup, ref = "15-24")
  )
```


## Fit the GEE Models

```{r}
fit_gee <- geeglm(
  health_bin ~ baseline_bin + txt + time + agegroup, id = id,
  data = df_gee, 
  family = binomial(link = "logit"),
  corstr = "unstructured"
)

summary(fit_gee)
```


## Get Odds Ratios

```{r}
coef_gee <- coef(summary(fit_gee))

tab_or_gee <- tibble(
  term = rownames(coef_gee),
  estimate = coef_gee[, "Estimate"],
  se = coef_gee[, "Std.err"],
  p_value = coef_gee[, "Pr(>|W|)"]
) |>
  mutate(
    OR = exp(estimate),
    lower_95 = exp(estimate - 1.96 * se),
    upper_95 = exp(estimate + 1.96 * se)
  )

tab_or_gee
```


* A logistic GEE model with an unstructured working correlation matrix was fit to the repeated follow-up observations, excluding the randomization visit. The outcome was participants’ self-rated health during follow-up, and the predictors were baseline health status, treatment group, follow-up visit and age group. This model accounts for the correlation among repeated observations from the same participant.  

* Baseline health was a strong predictor of follow-up self-rated health. After adjustment for treatment group, follow-up visit, and age group, participants who reported good health at baseline had significantly higher odds of reporting good health during follow-up than those who reported poor health at baseline (OR = 5.83, 95% CI: 2.07–16.45, p < 0.001).   

* Treatment group was also significantly associated with follow-up health. Participants in the intervention group had substantially higher odds of reporting good health during follow-up than those in the control group (OR = 6.75, 95% CI: 2.35–19.45, p < 0.001), after controlling for baseline health, follow-up visit, and age group.  

* There was no statistically significant evidence that self-rated health differed across follow-up visits. Relative to `time = 2`, the odds of reporting good health were not significantly different at `time = 3` (OR = 1.31, 95% CI: 0.67–2.56, p = 0.422) or `time = 4` (OR = 1.36, 95% CI: 0.57–3.21, p = 0.489).  

* Age group was partially associated with follow-up health. Compared with participants aged 15–24 years, those aged 25–34 years had significantly higher odds of reporting good health during follow-up (OR = 5.21, 95% CI: 1.87–14.48, p = 0.002). The estimated odds were also higher for participants aged 35 years and older, but this difference was not statistically significant (OR = 2.72, 95% CI: 0.46–16.07, p = 0.268).  

* Overall, these results suggest that follow-up self-rated health was strongly associated with baseline health and treatment assignment, while there was no clear evidence of a significant difference across follow-up visits. The results also suggest a higher likelihood of good self-rated health among participants aged 25–34 years compared with those aged 15–24 years.  


# Question c

```{r}
fit_glmm <- glmer(
  health_bin ~ baseline_bin + txt + time + agegroup + (1 | id),
  data = df_gee,
  family = binomial(link = "logit")
)

summary(fit_glmm)
```


```{r}
beta <- fixef(fit_glmm)
se <- sqrt(diag(vcov(fit_glmm)))

tab_or_glmm <- tibble(
  term = names(beta),
  estimate = beta,
  se = se,
  OR = exp(beta),
  lower_95 = exp(beta - 1.96 * se),
  upper_95 = exp(beta + 1.96 * se)
)

tab_or_glmm
```


* A logistic generalized linear mixed-effects model with a subject-specific random intercept was fit to the follow-up observations. The outcome was follow-up self-rated health, and the predictors were baseline health, treatment group, follow-up visit (`time`), and age group. The subject-specific random intercept was included to account for the correlation among repeated observations from the same participant and to allow individuals to differ in their underlying propensity to report good health.  

* Baseline health was a strong predictor of follow-up self-rated health. After controlling for treatment group, follow-up visit, age group, and the participant-specific random intercept, participants with good baseline health had substantially higher odds of reporting good health during follow-up than those with poor baseline health (OR = 16.58, 95% CI: 2.34–117.52, p = 0.0049). Treatment group was also significantly associated with follow-up health. Participants in the intervention group had much higher odds of reporting good health during follow-up than those in the control group (OR = 31.63, 95% CI: 3.72–268.90, p = 0.0016), conditional on the same random intercept and the other covariates.  

* There was no statistically significant evidence that follow-up self-rated health differed across visits. Relative to `time = 2`, the odds of reporting good health were not significantly different at `time = 3` (OR = 1.55, 95% CI: 0.52–4.64, p = 0.433) or `time = 4` (OR = 1.43, 95% CI: 0.42–4.82, p = 0.568). Age group was partially associated with the outcome. Compared with participants aged 15–24 years, those aged 25–34 years had significantly higher odds of reporting good health during follow-up (OR = 9.76, 95% CI: 1.31–72.74, p = 0.026), whereas the difference between participants aged 35+ years and those aged 15–24 years was not statistically significant (OR = 7.30, 95% CI: 0.47–112.61, p = 0.155).  

* The estimated random intercept variance was 5.87, indicating substantial between-subject heterogeneity in the propensity to report good health, even after adjusting for the observed predictors. This supports the inclusion of a subject-specific random intercept.  

* The interpretations of this model differ from those of the GEE model because the GLMM provides subject-specific (conditional) effects, whereas the GEE provides population-averaged (marginal) effects. Thus, the odds ratios from the mixed model describe how predictors are associated with the odds of good health for individuals with the same random intercept, while the GEE odds ratios describe the average association at the population level. In nonlinear models such as logistic regression, these two interpretations are not generally the same, and the subject-specific odds ratios from the GLMM are often further from 1 than the population-averaged odds ratios from the GEE. This pattern is seen here, where the estimated GLMM odds ratios are larger than the corresponding GEE odds ratios, although the overall substantive conclusions are similar across the two approaches.  























